---
title: "R Coding Exercise"
---

Placeholder file for the future R coding exercise.

```{r}
#adding in the dslabs package and other packages
library(dslabs)
library(dplyr)
library(ggplot2)
library(tidyverse)
library(here)
```

#dslabs is now in and datasets are available

```{r}
#next is to read help file for gapminder
help(gapminder)
```

```{r}
#get an overview of data structure and summary
str(gapminder)
summary(gapminder)
```

```{r}
#determine the type of object gapminder is
class(gapminder)
```

```{r}
#Assign african countries to new variable africadata
africadata <- subset(gapminder, continent == "Africa")
```

```{r}
#Checking the new variable
str(africadata)
summary(africadata)
```

```{r}
#taking object africadata and creating mortlife (infant mortality x life expectancy) using the dplyr package
mortlife <- africadata %>%
  select(infant_mortality,life_expectancy)
```

```{r}
#check the new object
str(mortlife)
summary(mortlife)
```

```{r}
#repeat the same process but create new object poplife (population x life expectancy) using dplyr
poplife <- africadata %>%
  select(population,life_expectancy)
```

```{r}
#check new object poplife
str(poplife)
summary(poplife)
```

```{r}
#plotting life expectancy as a function of infant mortality 
ggplot(mortlife, aes(x = life_expectancy, y = infant_mortality, color="blue")) + geom_point()
figure_file=here("coding-exercise","lifeexpectancyxinfantmortality.png")
ggsave(filename = figure_file)
```

```{r}
#plotting population as a function of life expectancy
# I used chatgpt to get the code for making the x variable on the log scale
ggplot(poplife, aes(x = population, y = life_expectancy, color="blue")) + geom_point() + scale_x_log10()
 figure_file=here("coding-exercise","populationxlifeexpectancy.png")
```

```{r}
#figuring out what years have missing infant mortality data
missing_data_years <- africadata %>%
  filter(is.na(infant_mortality)) %>%
  select(year) %>%
  distinct()
print(missing_data_years)
```

#### We see that data is missing from 1960-1981 and then again in 2016. Lets avoid those years and focus on 2000.

```{r}
#creating a new object for africadata, looking only at data from 2000
africadata2000 <- africadata %>%
  filter(year == 2000)
```

```{r}
str(africadata2000)
summary(africadata2000)
```

#### let's try remaking the plots from above using the data from 2000

```{r}
mortlife2 <- africadata2000 %>%
  select(infant_mortality,life_expectancy)
poplife2 <- africadata2000 %>%
  select(population,life_expectancy)
ggplot(mortlife2, aes(x = life_expectancy, y = infant_mortality, color="blue")) + geom_point()
figure_file=here("coding-exercise","lifeexpectancyxinfantmortality2.png")
ggsave(filename = figure_file)
ggplot(poplife2, aes(x = population, y = life_expectancy, color="blue")) + geom_point() + scale_x_log10()
 figure_file=here("coding-exercise","populationxlifeexpectancy2.png")


```

#### Now we will use the lm function to fit the model

```{r}
#fit1 is infant mortality as a predictor for life expectancy
fit1 <- lm(life_expectancy ~ infant_mortality, data=africadata2000)
summary(fit1)
```

```{r}
#fit2 is population size as a predictor for life expectancy
fit2 <- lm(life_expectancy ~ population, data=africadata2000)
summary(fit2)
```

#### Based on the p-value infant mortality rate is a better predictor for life expectancy than population size.

## **This following section is contributed by Vincent Nguyen**

This analysis begins with loading and inspecting the data set, us_contagious_diseases

```{r}
# load in the package dslabs
library(dslabs)

# load in the dataset, us_contagious_diseases
data("us_contagious_diseases")

# look at help file for data
help("us_contagious_diseases")

# get an overview of data structure
str(us_contagious_diseases)

# get a summary of the data
summary(us_contagious_diseases)

# determine the type of object data is
class(us_contagious_diseases)
```

This section of the analysis focuses on creating objects using the data set. The object, hepadata, is created to narrow the scope of the analysis to just Hepatitis A data. The objects, georgiadata and californiadata, are created to focus on Hepatitis A data in these states. Additionally, a new column is added to calculate incidence rates. Lastly, a dataframe is created, combining the data from both states. This is later used for another part of the analysis.

```{r}
# load tidyverse package
library(tidyverse)

# create new object only containing pertussis data
hepadata <- us_contagious_diseases %>%
  filter(disease == "Hepatitis A")

# create a new object containing data from only Georgia and create new incidence rate variable
georgiadata <- hepadata %>%
  filter(state == "Georgia") %>%
  mutate(incidence_rate = (count / population) * 100000)

# create a new object containing data from only California and create new incidence rate variable
californiadata <- hepadata %>%
  filter(state == "California")%>%
  mutate(incidence_rate = (count / population) * 100000)

# create a new object containing both for comparisions later

# first need to add new column so that each dataset indicates the state
georgiadata$state <- "Georgia"
californiadata$state <- "California"

# combine data sets into one
combined_data <- rbind(georgiadata, californiadata)

```

This section creates a graph outlining the change in incidence rate over time in the state of GA and CA.

```{r}
# load package ggplot2
library(ggplot2)

# create graph plotting year as x and incidence rate as y for Georgia
georgia_incidence_graph <- ggplot(georgiadata, aes(x = year, y = incidence_rate)) + geom_point(alpha = 0.6, color = "blue") + labs(
  title = "Incidence rate of Hep. A over time in Georgia",
  x = "Time",
  y = "Incidence (per 100,000" )

# display the newly created graph
print(georgia_incidence_graph)

# create graph plotting year as x and incidence rate as y for California
california_incidence_graph <- ggplot(californiadata, aes(x = year, y = incidence_rate)) + geom_point(alpha = 0.6, color = "blue") + labs(
  title = "Incidence rate of Hep. A over time in California",
  x = "Time",
  y = "Incidence (per 100,000" )

# display the newly created graph
print(california_incidence_graph)


```

This section creates a simple linear model to analyze the interaction of incidence rate over time across the two states.

```{r}
# linear model with interaction between year and state
lm_model <- lm(incidence_rate ~ year * state, data = combined_data)
summary(lm_model)

```

From these results, it can be concluded that Georgia has a significantly lower baseline incidence rate of Hepatitis A compared to California, and additionally, the change in incidence over time differs between the two states. The interaction term between year and state indicates that the change in incidence over time is significantly different between the two states. The incidence rate in California significantly decreases at a greater rate than Georgia indicating a possibly more effective response to Hepatitis A.
