---
title: "Moduel 4 Exercise"
---

### For this exercise I will be analyzing an Image. #The first step is to load in the necessary packages for exploring images. This will be magick and imager. Both packages can be used to manipulate the images properties such as its size, colors, effects, and even animating. Imager won't run on MacOS unless you install XQuartz first [here](https://www.xquartz.org/)

```{r}
library(tidyverse)
library(magick)
library(imager)
```

#### Now I will load in my images, my images are characters from a show so I will use their names

```{r}
jjk <- list(
  itadori = image_read("/Users/alexisgonzalez/Downloads/itadori.jpg"),
  nobara = image_read("/Users/alexisgonzalez/Downloads/nobara.jpg"),
  megumi = image_read("/Users/alexisgonzalez/Downloads/megumi.jpg"),
  nanami = image_read("/Users/alexisgonzalez/Downloads/nanami.jpg"),
  gojo = image_read("/Users/alexisgonzalez/Downloads/gojo.jpg"),
  sukuna = image_read("/Users/alexisgonzalez/Downloads/sukuna.jpg")
  )

```

#### The sizes of the images are different so I'm going to make them all the same 640x640

```{r}
resized_images <- lapply(jjk, function(img) {image_resize(img, "640x640" )})
```

#### Now I can combine all of these images and turn them into a gif

```{r}
gif <- image_animate(image_join(resized_images), fps = 5)
```

#### Now I view my gif, it should play in the window on the bottom left

```{r}
print(gif)
```

# 
